---
title: "Iteration_and_list columns"
output: github_document
date: "2024-10-31"
---

```{r setup}
library(tidyverse)
library(rvest)
```
## here are some lists:
```{r}
l = list(
  vec_numeric = 1:4,
  unif_sample = runif(100),
  mat = matrix(1:8, nrow = 2, ncol = 4, byrow = TRUE),
  summary = summary(rnorm(1000))
)

l
```
access lists
```{r}
l$mat

l[["mat"]]
```

make a list hopefully a bit more useful
```{r}
list_norm = 
  list(
    a = rnorm(20, 0, 5),
    b = rnorm(20, 4, 5),
    c = rnorm(20, 0, 10),
    d = rnorm(20, 4, 10)
  )

list_norm$b
```
reuse function we wrote last time
```{r}
mean_and_sd = function(x) {
  
  mean_x = mean(x)
  sd_x = sd(x)
  
  out_df = 
    tibble(
      mean = mean_x,
      sd = sd_x
    )
  return(out_df)
}

```
use the function to take mean and sd for all normal samples
```{r}
mean_and_sd(list_norm[["a"]])
mean_and_sd(list_norm[["b"]])
mean_and_sd(list_norm[["c"]])
mean_and_sd(list_norm[["d"]])
```
create output list from a for loop:
a vector of type list and length 4
```{r}
output = vector("list", length = 4)

for (i in 1:4) {
  output[[i]] = mean_and_sd(list_norm[[i]])
}

output
```

## do the same thing with `map` instead
here is the list that I want, please apply the function mean_and_sd
```{r}
output = map(list_norm, mean_and_sd)

output
```
let's do a couple other things

combine lists
```{r}
output = map(list_norm, IQR)

output = map_dbl(list_norm, IQR)
```

combine a list of dataframes
```{r}
output = map_dfr(list_norm, mean_and_sd)

#equivalent to
output = map(list_norm, mean_and_sd) %>% 
  bind_rows()
```

## LIST COLUMNS!!
you can do normal df operations
```{r}
listcol_df =
  tibble(
    name = c("a", "b", "c", "d"),
    samp = list_norm
  )

listcol_df %>% 
  filter(name %in% c("a", "b"))

listcol_df %>% 
  select(-samp)
```

pull out a sample column
```{r}
listcol_df[["samp"]][["a"]]
```
compute mean and sd
```{r}
mean_and_sd(listcol_df[["samp"]][["a"]])

#do the same thing as before: map over a list and compute mean and sd
map(listcol_df[["samp"]], mean_and_sd)
```
ADD A LIST COLUMN!!!
```{r}
listcol_df %>% 
  mutate(
    output = map(samp, mean_and_sd),
    iqr = map_dbl(samp, IQR))

listcol_df %>% 
  mutate(
    output = map(samp, mean_and_sd),
    iqr = map_dbl(samp, IQR)) %>% 
  select(-samp) %>% 
  unnest(output)
```
### NSDUH
goal: have a df that has everything

new version of the nsduh function
```{r}
nsduh_import = function(html, nth_table) {
  
  out_df = 
    html %>% 
    html_table() %>% 
    nth(nth_table) %>% 
    slice(-1) %>% 
    select(-contains("P Value"))
  
  return(out_df)
}
```
we need to import the html, extract correct tables

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)
```

```{r}
nsduh_import(html = nsduh_html, nth_table = 1)
nsduh_import(html = nsduh_html, nth_table = 4)
nsduh_import(html = nsduh_html, nth_table = 5)
```
```{r}
nsduh_listcol_df =
  tibble(
    drug = c("marj", "cocaine", "heroin"),
    nth_table = c(1, 4, 5),
  ) %>% 
  mutate(table = map(nth_table, nsduh_import, html = nsduh_html)) %>% 
#html is the only argument left, add as additional argument
  unnest(table)
```

Now we have a well-organized df
```{r}
nsduh_listcol_df %>% 
  filter(State == "New York")
```
## Operations on nested data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```
create a list column.
```{r}
weather_nest= 
  weather_df %>% 
  nest(data = date:tmin)
```

```{r}
weather_nest[["data"]][[1]]
```

regressing tmax on tmin.
```{r}
lm(tmax ~ tmin, data = weather_nest[["data"]][[1]])
lm(tmax ~ tmin, data = weather_nest[["data"]][[2]])
lm(tmax ~ tmin, data = weather_nest[["data"]][[3]])
```
```{r}
#here is my data, call it x, then plug it into the lm function
weather_nest %>% 
  mutate(model_fit = map(data, \(x) lm(tmax~tmin, data = x)))
```

